### 2024-01-05 sections/1-introduction.tex
incorporate some of Carl's suggestions
addressing carl's comments

Researchers are collecting data at an unprecedented scale.
In many fields, the sizes of datasets are growing exponentially, and this increase in the rate of data collection outpaces the rate of improvements in computing performance as predicted by Moore's Law~\cite{brescia2012extracting}.
This indicates that computer architecture will not ``catch up'' to computational needs in the near future.
Often dubbed ``the Big Data explosion,'' this phenomenon has created a need for better algorithms for analyzing such datasets.

@@

InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} is a method for approximate $k$-NN search. 
The data are clustered into high-dimensional Voronoi cells, and whichever cell the query point into is then searched exhaustively.
The number of cells used is governed by the $n_{list}$ parameter. 
Increasing this parameter decreases the number of points being exhaustively searched, so it improves speed at the cost of accuracy.
To mitigate accuracy issues caused by a query point falling near a cell boundary, the algorithm has a tunable parameter $n_{probe}$, which specifies the number of additional adjacent or nearby cells to search.

>>>>

Researchers are collecting data at an unprecedented scale.
In many fields, the sizes of datasets are growing exponentially, and this increase in the rate of data collection outpaces the rate of improvements in computing performance as predicted by Moore's Law~\cite{kahn2011future}.
This indicates that computer architecture will not ``catch up'' to computational needs in the near future.
Often dubbed ``the Big Data explosion,'' this phenomenon has created a need for better algorithms for analyzing such datasets.

@@

InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} is a method for approximate $k$-NN search. 
The data are clustered into high-dimensional Voronoi cells, and whichever cell the query point falls into is then searched exhaustively.
The number of cells used is governed by the $n_{list}$ parameter. 
Increasing this parameter decreases the number of points being exhaustively searched, so it improves speed at the cost of accuracy.
To mitigate accuracy issues caused by a query point falling near a cell boundary, the algorithm has a tunable parameter $n_{probe}$, which specifies the number of additional adjacent or nearby cells to search.

### 2024-01-05 sections/2-methods.tex
addressing carl's comments

\begin{gather}
    \mathcal{O}
    \Bigg(
        \underbrace{
            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
        }_{\textrm{tree-search}}
        \ + \ 
        \underbrace{
            \overbrace{ \big| B_X(q, \rho) \big|}^{\textrm{output size}}
            \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ d}^{\textrm{scaling factor}}
        }_{\textrm{leaf search}}
    \Bigg)
    \label{eq:methods:rnn-search-complexity}
\end{gather}

@@

where $\mu$ is the harmonic mean of the LFD of the clusters in $S$.
We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD.
We allow a maximum increase of 2 to ensure that we do not increase the radius too quickly in any single iteration.

@@

Using this to simplify the term for leaf search in Equation~\ref{eq:methods:rnn-search-complexity}, we get

@@

\begin{gather}
    \mathcal{O}
    \Bigg(
        \underbrace{
            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
        }_{\textrm{tree-search}}
        \ + \ 
        \underbrace{
            \overbrace{k}^{\textrm{output size}}
            \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
        }_{\textrm{leaf search}}
    \Bigg)
    \label{eq:methods:repeated-rnn-complexity}
\end{gather}

>>>>

\begin{gather}
    \mathcal{O}
    \Bigg(
        \underbrace{
            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
        }_{\textrm{tree-search}}
        \ + \ 
        \underbrace{
            \overbrace{ \big| B_X(q, \rho) \big|}^{\textrm{output size}}
            \overbrace{ \left( \frac{\rho + 2 \cdot \hat{r}}{ \rho} \right) ^ d}^{\textrm{scaling factor}}
        }_{\textrm{leaf-search}}
    \Bigg)
    \label{eq:methods:rnn-search-complexity}
\end{gather}

@@

where $\mu$ is the harmonic mean of the LFD of the clusters in $S$.
We use the harmonic mean to ensure that $\mu$ is not dominated by outlier clusters with very high LFD.
We cap the radial increase at 2 to ensure that we do not increase the radius too quickly in any single iteration.

@@

Using this to simplify the term for leaf-search in Equation~\ref{eq:methods:rnn-search-complexity}, we get

@@

\begin{gather}
    \mathcal{O}
    \Bigg(
        \underbrace{
            \log~\overbrace{\mathcal{N}_{\hat{r}}(X)}^{\textrm{metric entropy}}
        }_{\textrm{tree-search}}
        \ + \ 
        \underbrace{
            \overbrace{k}^{\textrm{output size}}
            \overbrace{\bigg( 1 + 2 \cdot \Big( \frac{\hat{|C|}}{k} \Big) ^ {d^{-1}} \bigg)^d}^{\textrm{scaling factor}}
        }_{\textrm{leaf-search}}
    \Bigg)
    \label{eq:methods:repeated-rnn-complexity}
\end{gather}

### 2024-01-05 sections/4-results.tex
addressing carl's comments

We also highlight some differences in indexing time between different datasets.
With Fashion-Mnist, as shown in Figure~\ref{fig:results:fashion-mnist-indexing}, we observe that the indexing time for CAKES is lower than that of FAISS-IVF for all cardinalities.
With Glove-25 (see Figure~\ref{fig:results:glove-25-indexing}), however, at cardinalities greater than $10^7$, FAISS-IVF has lower indexing time than CAKES.
With Sift, CAKEs's indexing time is faster than that of FAISS-IVF until a cardinality of nearly $10^8$, and with the Random dataset, we observe that CAKES has faster indexing time than FAISS-IVF until a cardinality of nearly $10^7$.

@@

For Silva and RadioML, we benchmarked only CAKES's algorithms because none HNSW, ANNOY or FAISS support neither the required distance functions nor, in the case of RadioML, complex-valued data.
Due to the massive sizes of these datasets, we took random sub-samples of the dataset with lower cardinalities for our benchmarks, rather than augmented versions of the dataset.
With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput seems to linearly decrease as cardinality increases, but that it seems to begin levelling off near cardinality $10^5$.
Until cardinality near $10^4$, Depth-First Sieve is the fastest CAKES algorithm, but for larger cardinalities, Repeated $\rho$-NN is the fastest CAKES algorithm.
For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly the three CAKES algorithms exhibit nearly indistinguishable performance for all cardinalities.
In particular, throughput of CAKES algorithms is identical within three significant figures for all cardinalities.

>>>>

We also highlight some differences in indexing time between different datasets.
With Fashion-Mnist, as shown in Figure~\ref{fig:results:fashion-mnist-indexing}, we observe that the indexing time for CAKES is lower than that of FAISS-IVF for all cardinalities.
With Glove-25 (see Figure~\ref{fig:results:glove-25-indexing}), however, at cardinalities greater than $10^7$, FAISS-IVF has lower indexing time than CAKES.
With Sift, CAKES's indexing time is faster than that of FAISS-IVF until a cardinality of nearly $10^8$, and with the Random dataset, we observe that CAKES has faster indexing time than FAISS-IVF until a cardinality of nearly $10^7$.

@@

For Silva and RadioML, we benchmarked only CAKES's algorithms because HNSW, ANNOY and FAISS do not support neither the required distance functions nor, in the case of RadioML, complex-valued data.
Due to the massive sizes of these datasets, we took random sub-samples of the dataset with lower cardinalities for our benchmarks, rather than augmented versions of the dataset.
With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput seems to linearly decrease as cardinality increases, but that it seems to begin levelling off near cardinality $10^5$.
Until cardinality near $10^4$, Depth-First Sieve is the fastest CAKES algorithm, but for larger cardinalities, Repeated $\rho$-NN is the fastest CAKES algorithm.
For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly the three CAKES algorithms exhibit nearly indistinguishable performance for all cardinalities.
In particular, throughput of CAKES's algorithms is identical within three significant figures for all cardinalities.
